---
title: "simplefunctions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{simplefunctions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Simplefunctions is an R package consisting of four functions, that are really basic and quite specific.The functions are:

* Function 1: `get_date()`
* Function 2: `disease_count()`
* Function 3: `plot_cases()`
* Function 4: `load_image()`

These functions will be used with examples below.

The package also comes with 2 datasets named `flu_cases` and `dengue_cases`. With `flu_cases` showing the estimated flu cases for different countries per year, and `dengue_cases` showing the same thing, but for estimated dengue cases. 
These datasets are documented, and documentation can be read by using `?flu_cases` and `?dengue_cases`.


We will start by loading in our package and both datasets, then using each function to show examples of how they work.

```{r loading package and datasets }
devtools::load_all()
library(simplefunctions)

data("flu_cases")
data("dengue_cases")
```


## Function 1: `get_date()`
When a dataframe contains a column with dates in yyyy/mm/dd format, this function can be used to get only the years, months or days. I have used the code of this function to get the years for my datasets. In the chunk below I'm showing how it would work on a sample dataframe.
```{r get_date() example}

# Creating dataframe with random years in yyyy/mm/dd format
dataframe <- data.frame(date = c(sample(seq(as.Date('1999/01/01'), as.Date('2023/01/01'), by="year"), 12)))
dataframe


# Using get_date() to obtain the years.
dataframe$years <- get_date(dataframe$date,dmy = "year")
dataframe


# Using get_date() to obtain the months.
dataframe$months <- get_date(dataframe$date,dmy = "month")
dataframe

# Using get_date() to obtain the days.
dataframe$days <- get_date(dataframe$date,dmy = "day")
dataframe
```



## Function 2: `disease_count()`
This function counts the total amount of cases of dengue or flu in a country throughout the years. This function is fairly specific for the 2 datasets in this package, so these will be used as examples.

```{r disease_count() example}
# Total amount of flu cases from 2002-2015 in Belgium. 
disease_count(df = flu_cases, country_name = "Belgium", disease = "flu")

# Total amount of dengue cases from 2002-2015 in Brazil.
disease_count(df = dengue_cases, country_name = "Brazil", disease = "dengue")

```
## Function 3: `plot_cases()`

This function plots a line graph of dengue/flu cases throughout the years in different countries. It uses all the countries in the dataframe, so it is best used after filtering for only the countries you want. 
```{r plot_cases() example, fig.width= 7}

# Filtering only Argentina, Bolivia and Brazil for dengue_cases
argbolbra_dengue <- dengue_cases %>% filter(country == "Argentina" | country == "Bolivia" | country == "Brazil")

# Plotting dengue cases for Argentina, Bolivia and Brazil from 2002-2015
plot_cases(df = argbolbra_dengue, disease = "dengue",plot_title = "Dengue cases Argentina, Bolivia and Brazil from 2002-2015")

# Filtering only Argentina, Bolivia and Brazil for flu_cases
argbolbra_flu <- flu_cases %>% filter(country == "Argentina" | country == "Bolivia" | country == "Brazil")

# Plotting flu cases for Argentina, Bolivia and Brazil from 2002-2015
plot_cases(df = argbolbra_flu, disease = "flu",plot_title = "Flu cases Argentina, Bolivia and Brazil from 2002-2015")
```



## Function 4: `load_image()`
This function is not really related to previous functions. This one's just for loading an image by giving the path to the file.

```{r load_image() example, fig.height=5, fig.width=5}
# Loading a picture a folder structure from a course I followed.
load_image("inst/DAUR2_fs.png")
```



